---
title: "DATA TABLES in R"
author: "Vaishali Chaudhuri"
date: "May 23, 2015"
output: ioslides_presentation
---

## How to Use Data Table Package in R

This data.table R lecture explains the basics of the DT[i, j, by] command which is core to the data.table package. If you want to learn more on the data.table package see <https://github.com/Rdatatable/data.table/wiki>. Slides Adapted from <http://blog.datacamp.com/data-table-r-tutorial/>



## What are the advantages of using Data Table Package

- With large datasets a data.frame can be limiting: the time it takes to do certain things is just too long.
- Data.table solves this for you by reducing computing time. 
- It also makes it easier to do more with less typing. 
- So you will not only be reducing computing time, but programming time as well.


## Basic arguments to DataTable

The DT[i,j,by] command has three parts: i, j and by. If you think in SQL terminology, 
the i corresponds to WHERE, 
j to SELECT and 
by to GROUP BY. 
We talk about the command by saying “Take DT, subset the rows using ‘i’, then calculate ‘j’ grouped by ‘by’.


## An example

In a simple example and using the hflights dataset (so  all the examples) can be demonstarted:

```{r eval=FALSE}
library(hflights)
library(data.table)
DT &lt;- as.data.table(hflights)
DT[Month==10,mean(na.omit(AirTime)),
                  by=UniqueCarrier]
UniqueCarrier               V1
1:         AA         68.76471
2:         AS         255.29032
3:         B6         176.93548
4:         CO         141.52861
```

## What is the 'j' part

- Here we subsetted the data table to keep only the rows of the 10th Month of the year, 
- Calculated the average AirTime of the planes that actually flew (that’s why na.omit() is used, 
- Cancelled flights don’t have a value for their AirTime) and then grouped the results by their Carrier.
- We can see for example that AA (American Airlines) has a very short average AirTime compared to AS (Alaska Airlines). 
- Notice that R base functions can be used in the j part? 


## What is the 'i' part

The ‘i’ part is used for subsetting on rows, just like in a data frame.

```{r eval=FALSE}

DT[2:5]
#selects the second to the fifth row of DT
Year Month DayofMonth DayOfWeek DepTime ArrTime
1: 2011 1 2 7 1401 1501 AA 428 N557AA 60 45
2: 2011 1 3 1 1352 1502 AA 428 N541AA 70 48
3: 2011 1 4 2 1403 1513 AA 428 N403AA 70 39
4: 2011 1 5 3 1405 1507 AA 428 N492AA 62 44
ArrDelay DepDelay Origin Dest Distance TaxiIn    
1: -9 1 IAH DFW 224 6 9 0 0
2: -8 -8 IAH DFW 224 5 17 0 0
3: 3 3 IAH DFW 224 9 22 0 0
4: -3 5 IAH DFW 224 9 9 0 0
```

## Column names can also be usedas they are evaluated in the scope of DT
```{r eval=FALSE}
DT[UniqueCarrier==&quot;AA&quot;]
#Returns all those rows where the Carrier is
American Airlines
Year Month DayofMonth DayOfWeek DepTime ArrTime
UniqueCarrier FlightNum TailNum
ActualElapsedTime
```

## Quirky and useful:
- Don’t have to use a comma for subsetting rows in a data table. 
  we have to specify DF[2:5,]. 
- when subsetting rows you can also use the symbol .N in the DT[…] command, 
  which is the number of rows or the last row. You can use it for selecting the 
  last row or an offset from it.
```{r eval=FALSE}
T[.N-1]
#Returns the penultimate row of DT
Year Month DayofMonth DayOfWeek DepTime ArrTime
UniqueCarrier FlightNum TailNum ActualElapsedTime AirTime
1: 2011 12 6 2 656 812 WN 621 N727SW 76 64
ArrDelay DepDelay Origin Dest Distance
TaxiIn TaxiOut Cancelled CancellationCode
Diverted
1: -13 -4 HOU TUL 453 3 9 0 0
```

## The j part

The ‘j’ part is used to select columns and do stuff with them. 
And stuff can really mean anything. All kinds of functions can be used, 
which is a strong point of the data.table package.

```{r eval=FALSE}
DT[, mean(na.omit(ArrDelay))]
[1] 7.094334
```

##
- Notice that the ‘i’ part is left blank, and the first thing in the brackets is a comma. 
- This might seem counterintuitive at first. However, this simply means that 
  we do not subset on any rows, so all rows are selected. 
- In the ‘j’ part, the average delay on arrival of all flights is calculated. 
- When selecting several columns and doing stuff with them in the ‘j’ part, 
  you need to use the ‘.()’ notation. This notation is actually just an alias to ‘list()’. 
- It returns a data table, whereas not using ‘.()’ only returns a vector, as shown above.

## 
```{r eval=FALSE}
DT[, .(mean(na.omit(DepDelay)),
mean(na.omit(ArrDelay)))]
         V1       V2
1: 9.444951 7.094334

```

##

Another useful feature which requires the ‘.()’ notation allows you to rename columns inside the DT[…] command.

```{r eval=FALSE}
DT[, .(Avg_ArrDelay =mean(na.omit(ArrDelay)))]
Avg_ArrDelay
1: 7.094334
DT[, .(Avg_DepDelay = mean(na.omit(DepDelay)),
    avg_ArrDelay = mean(na.omit(ArrDelay)))]
Avg_DepDelay Avg_ArrDelay
1:  9.444951     7.094334
```
Of course, new column names are not obligatory.

## 
Combining the above about ‘i’ and ‘j’ gives:

```{r eval=FALSE}
DT[UniqueCarrier==&quot;AA&quot;, .(Avg_DepDelay =
mean(na.omit(DepDelay)),
Avg_ArrDelay = mean(na.omit(ArrDelay)),
plot(DepTime,DepDelay,ylim=c(-15,200)),
abline(h=0))]
Avg_DepDelay Avg_ArrDelay   V3  V4
1: 6.390144     0.8917558 NULL NULL
```

##
Here we took DT, selected all rows where the carrier was AA in the ‘i’ part, 
calculated the average delay on departure and on arrival, 
and plotted the time of departure against the delay on departure in the ‘j’ part.

To recap, the ‘j’ part is used to do calculations on columns specified in that part.
As the columns of a data table are seen as variables, and the parts of ‘j’ 
are evaluated as expressions, virtually anything can be done in the ‘j’ part. 
This significantly shortens  programming time.

## The 'by' part

- The final section of this data.table R tutorial focuses on the ‘by’ part. 
- The ‘by’ part is used when we want to calculate the ‘j’ part grouped by a specific 
variable (or a manipulation of that variable). - - You will see that the ‘j’ expression is repeated for each ‘by’ group. 
It is simple to use: you just specify the column you want to group by in the ‘by’ argument.

## 
```{r eval=FALSE}
DT[,mean(na.omit(DepDelay)),by=Origin]
Origin       V1
1: IAH 8.436951
2: HOU 12.837873
```

##
- Here we calculated the average delay before departure, but grouped by where the plane is coming from. 
- It seems that flights departing from HOU have a larger average delay than those leaving from IAH.
- Just as with the ‘j’ part, you can do a lot of stuff in the ‘by’ part. 
- Functions can be used in the ‘by’ part so that results of the operations done in the 
‘j’ part are grouped by something we specified in the DT[…] command. 
- Using functions inside DT[…] makes that one line very powerful. 
- Likewise, the ‘.()’ notation needs to be used when using several columns in the ‘by’ part.

##
```{r eval=FALSE}

DT[,.(Avg_DepDelay_byWeekdays =
   mean(na.omit(DepDelay))),
   by=.(Origin,Weekdays = DayOfWeek&lt;6)]
Origin Weekdays Avg_DepDelay_byWeekdays
1: IAH FALSE     8.286543
2: IAH TRUE      8.492484
3: HOU FALSE    10.965384
4: HOU TRUE     13.433994
```

##
Here, the average delay before departure of all planes (no subsetting in the ‘i’ part, so all rows are selected) 
was calculated first, and grouped secondly, first by origin of the plane and then by weekday. 
Weekdays is False in the weekends. It appears that the average delay before departure 
was larger when the plane left from HOU than from IAH, and surprisingly the delays were smaller in the weekends
Putting it all together a typical DT[i,j,by] command gives:
```{r eval=FALSE}
DT[UniqueCarrier==&quot;DL&quot;, .(Avg_DepDelay =
mean(na.omit(DepDelay)),
Avg_ArrDelay = mean(na.omit(ArrDelay)),
Compensation = mean(na.omit(ArrDelay - DepDelay))),
by = .(Origin, Weekdays = DayOfWeek&amp;lt;6)]
Origin Weekdays Avg_DepDelay Avg_ArrDelay Compensation
1: IAH FALSE    8.979730     4.116751     -4.825719
2: HOU FALSE    7.120000     2.656566     -4.555556
3: IAH TRUE     9.270948     6.281941     -2.836609
4: HOU TRUE     11.631387    10.406593    -1.278388
```


## Conclusion
- Here the subset of planes flewn by Delta Air Lines (selected in ‘i’) 
was grouped by their origin and by Weekdays (in ‘by’).
- The time that was compensated in air was also calculated (in ‘j’). 
It appears that in the weekends, irrespective of the plane was coming from IAH or HOU, 
the time compensated while in air (thus by flying faster) is bigger.