---
title: "Lecture 1b"
author: "Vaishali Chaudhuri"
date: "April 3, 2015"
output: slidy_presentation
---

## What is  Github

- Github is like a mini filesystem with  powerful tools of version control built in

- Git thinks of its data more like a set of snapshots of a miniature filesystem.

- Most operations in Git only need local files and resources to operate 

- When you do actions in Git, nearly all of them only add data to the Git database. It is hard to  erase data in any way

## The Three States

Git has three main states that your files can reside in: 

- Committed means that the data is safely stored in your local database 

- Modified means that you have changed the file but have not committed it to your database yet

- Staged means that you have marked a modified file in its current version to go into your next commit snapshot


This leads us to the three main sections of a Git project: the Git directory, the working directory, and the staging area.


## How to install git on computer 

- In Mac
    - GitHub has a Mac client! You can use it without ever touching the command line.

- In Windows
  - Download and install the latest version of Git.
  - On your computer, open the Terminal application.
  - Tell Git your name so your commits will be properly labeled. Type everything        after the $ here:

```
$ git config --global user.name "YOUR NAME"

```
-Associate  the email address that will be associated with your Git commits. Keep your email address setting as private.

```
$ git config --global user.email "YOUR EMAIL ADDRESS"

```

## Next steps: Authenticating with GitHub from Git

- When you connect to a GitHub repository from Git, you'll need to authenticate with GitHub using either HTTPS or SSH.

- Connecting over HTTPS (recommended)

  - If you clone with HTTPS, you can cache your GitHub password in Git using a credential helper.

- Connecting over SSH

  _ If you clone with SSH, you must generate SSH keys on each computer you use to push or pull from GitHub.

## How to create a repo


On GitHub, one can store all kinds of projects in repositories.

-Personal repositories belong to user accounts,  after you've signed up for GitHub, create your first repository!

-   In the upper-right corner of you account page click- New repository
-   Create a short, memorable name for your repository. For example, "hello-world".

- Choose between creating a public or private repository.

- Public repositories are a great choice and visible to any user on GitHub, so you can benefit from a collaborative community.
- Private repositories are only available for paid accounts only available to you, the repository owner.


## Commit your first change

A commit is like a snapshot of all the files in your project at a particular point in time.

 - README files are a great place to describe your project in more detail, or add some documentation such as how to install or use your project. 
 - The contents of your README file are automatically shown on the front page of your repository.



Let's commit a change to the README file.

## Commit changes (contd.)

- In your repository's list of files, click README.md.
- Above the file's content, click .
- On the Edit file tab , type some information about yourself.
- Above the new content, click Preview changes.
- File preview viewReview the changes you made to the file. You'll see the new content in green.
- At the bottom of the page, under "Commit changes," type a short, meaningful commit message that describes the change you made to the file.
- Click Commit changes.

Congratulations! You have now created a repository, including a README file, and created your first commit on GitHub.

## Introducing concepts in LaTeX document

- Why write in LaTeX?
    - Pros (why prepare presentations using LaTeX):
        - The software for preparing the presentation is free.
        - The outcome is platform-independent: you can display PDF documents pretty     much with any OS.
        - You get the typographic advantages of LaTeX, especially in typesetting of text and mathematics.
    - Cons (why you might prefer MS Powerpoint/Word):
        - If you have no or very little math, the advantages of LaTeX becomes less critical.
        - Some aspects of animations may be admittedly harder to create with LaTeX 
        - It is (yet) impossible to embed movies or sound in the PDF presentation 
  
## Write a document with  in LaTeX using scientific concepts

Typesetting mathematics is one of LaTeX's greatest strengths. 

- If your document requires only a few simple mathematical formulas, plain LaTeX has most of the tools that you will need. 

- If you are writing a scientific document that contains numerous complicated formulas, the 'amsmath' package introduces several new commands that are more powerful and flexible than the ones provided by LaTeX. 


## Literate Calculation Exercise

- Use simple $\LaTeX$ formatting to typeset a mathematical equation in an RMarkdown document.

- check out (calcCheck1.RMD/pdf) as an example of Latex scientific document for homework:
https://github.com/rmhorton/statprog/tree/master/week_01_teach_yourself_R

- check out how to write more Mathematical font at:

http://en.wikibooks.org/wiki/LaTeX/Mathematics


## What would this function do

What does this function do?

```{r, eval=FALSE}
function(N) {
  x <- runif(N) 
  y <- runif(N)
  4 * sum( (x^2 + y^2) < 1 )/N
}
```
* How would you give this function  a name?

## Solution to the previous problem 

- start of by making two uniformly distributed random numbers x and y
- ((x^2 + y^2) < 1 ) -This expression  depicts  the square of each x and y  such that the square of them is less than 1 
- Geometrically thinking on a graph this represents  random points on the graph whose sum of squares  is going to be less than 1 . 
- (sqrt(x^2 +y^2) < 1)-The square root of the (sum( (x^2 + y^2)) will give distance of x and y vector from point zero.
- ((x^2 + y^2) < 1 ). This gives a logical vector
_ sum of logical vector coerces a logical vector into true and false and in this case tells if the distance of a point from zero is less than 1 or not



## Soln to the problem continued...
- Multiplication by 4 gives the representation in all the four quadrants from distance 0,0 as each quadrant is a mirror image of each other
- Division by N gives the ratio of points in a unitcircle inside a unit square
$$ r = 1 $$
$$ A_{circle} = 4* \pi r^2  $$
$$ A_{square} = (2 r) ^ 2 = 4r^2 $$
$$  4* \pi r^2 / 4r^2 = \pi $$
Therefore this function can be named 
 - How to approximately estimate value of Pi 



## Homework Exercise


```{r, eval=FALSE}
function(N) {
  x <- runif(N)
  y <- runif(N)
  4 * sum( (x^2 + y^2) < 1 )/N
}
```

- What is the type of the parameter `N`? (hint: see `?runif`)
- How can you call this function to see what happens?
- What is the type of the argument to `sum()`?


## Homework Exercise Review

What will this code show?

```{r throwing_darts, height=5, width=5}
N <- 10000
x <- runif(N, min=-1.0, max=1.0)
y <- runif(N, min=-1.0, max=1.0)
plot(x, y, pch=16, col=ifelse(x^2 + y^2<1, "red", "blue"))
```

## Homework Exercise Review

What will this code show?

```{r darts_one_quad, height=5, width=5}
N <- 10000
x <- runif(N)
y <- runif(N)
plot(x, y, pch=16, col=ifelse(x^2 + y^2<1, "red", "blue"))
```


## Homework Exercise Review

What will this code show?

```{r review_estimate_pi}
estimate_pi <- function(N) {
    x <- runif(N)
    y <- runif(N)
  4 * sum( (x^2 + y^2) < 1 )/N
}
estimate_pi(100000)
```
